---
title: "Demand Analysis for Company X"
format: html
date: "2015-01-15"
date-modified: "2024-11-25"
---

# Task

TASKS (do not delete formulas+ add your comments if needed)

1. CE, Nordic, Other areas total 2015 sales estimation: 500 000pcs. However, we expect (compared to 2014 sales):
A. Scenario 1
a. 5% increase of the CE share 
b. 2% decrease of Other share
c. 3% decrease of Nordic share

B. Scenario 2
a. 10% increase of the CE share
b. 1% decrease of Other share
c.  9% decrease of Nordic share
Regarding the product mix, we expect 10% growth of prodgroup 134 in all areas

Using the information in this workbook, please calculate estimation 2015 for Scenario 1 and 2 by items and sales areas as shown below. Please make it so that we could choose Scenario 1 or 2. All 0 values should be blank. Formatting is up to you

2. Please make a comparison between Forecast 2015 Scenario 1 and sales 2014 both on total and item level (no sales area split is needed). Use charts if you want


# Conclusion from 2013

"Task 1
Scenario 1 for three market areas is calculated using rate of change information given in the assignment. For Scenario 2, I created a custom filter (a combination of Excels IF and OR functions) for identifying “prodgroup 134” from different market areas (see Sales 2014 sheet). 

For scenario selection I created two “option buttons”, this approach utilizes Excels IF functions for viewing scenarios in different sales areas. I also calculated Prodgroup 1 and item 1 in the tables, since Progroup 1 only contains item 1 and is sold only in the CE area, cells for other market areas remain blank."

"Task 2
I used Excels SUMIF functions to combine possible duplicate items on sales 2014 and 2015 list. For comparison, I created a chart for items and total sales.
"

# libraries

```{r}
library(tidyverse)
library(readxl)
```


# data import

```{r}
filename <- "Excel test.xlsx"
sales2014 <- read_excel(filename, sheet = 2, range = cell_cols("A:C")) |> janitor::clean_names()

mapping <- read_excel(filename, sheet = 3) |> janitor::clean_names()
```



```{r}
df <- sales2014 |> full_join(mapping)
```


# Scenarios

## Scenario 1

1. CE, Nordic, Other areas total 2015 sales estimation: 500 000pcs. However, we expect (compared to 2014 sales):
A. Scenario 1
a. 5% increase of the CE share 
b. 2% decrease of Other share
c. 3% decrease of Nordic share


```{r}
df
yearly_sales_2014 <- df |>
  group_by(sales_area) |>
  summarise(total_sales_2014 = sum(sales_2014))
yearly_sales_2014


yearly_sales_scenario_1 <- yearly_sales_2014 |> mutate(
  estimate_2015_scenario_1 = case_when(
    sales_area == "CE" ~ total_sales_2014 * 1.05,
    sales_area == "Other" ~ total_sales_2014 * 0.98,
    sales_area == "Nordic" ~ total_sales_2014 * 0.97
  )
)

yearly_sales_scenario_1
```


## Scenario 2

a. 10% increase of the CE share
b. 1% decrease of Other share
c.  9% decrease of Nordic share
Regarding the product mix, we expect 10% growth of prodgroup 134 in all areas


```{r}
yearly_sales_prep_scenario2 <- df |>
  mutate(adjusted_sales_2015_prep = if_else(product_group == "prodgroup 134", sales_2014 * 1.10, sales_2014)) |>
  group_by(sales_area) |>
  summarise(total_sales_prep = sum(adjusted_sales_2015_prep))
yearly_sales_prep_scenario2

yearly_sales_scenatio_2 <- yearly_sales_prep_scenario2 |> transmute(
  estimate_2015_scenario_2 = case_when(
    sales_area == "CE" ~ total_sales_prep * 1.10,
    sales_area == "Other" ~ total_sales_prep * 0.99,
    sales_area == "Nordic" ~ total_sales_prep * 0.91
  ), sales_area
)
yearly_sales_scenatio_2
res <- yearly_sales_scenario_1 |> full_join(yearly_sales_scenatio_2, by = join_by("sales_area"))
```

## visualization
visualize bar chart res
```{r}
res |>
  pivot_longer(cols = c(estimate_2015_scenario_1, estimate_2015_scenario_2), names_to = "scenario", values_to = "sales") |>
  ggplot(aes(x = sales_area, y = sales, fill = scenario)) +
  geom_col(position = "dodge") +
  labs(title = "Scenario 1 and 2", x = "Sales area", y = "Sales") +
  # scale y to eur value
  scale_y_continuous(labels = scales::label_currency(prefix = "EUR ")) +
  theme_minimal()
```
